### 2014-03-27 paper/conclusion.tex
paper workin

\section{Conclusion}
While today's 3D printers are not yet able to fabricate active components in-place, we suggest that 3D printed interactive devices can be created with redirection of active input and output via tubes.  In addition to describing the design space of tubes and connecting related works created through both traditional processes and rapid prototyping methods, we also presented a design tool for the creation of tubes inside arbitrary 3D models.  We fabricated several example objects using this tool to explore new points in the design space.

We are heartened by our success: it opens new interactive device prototyping opportunities both for makers with consumer-grade 3D printers and research and industrial labs with higher-end machines.  We are excited to release our tool into the wild and see what people will build.

>>>>

\section{Conclusion}

While today's 3D printers are not yet able to fabricate active components in-place, we suggest that 3D printed interactive devices can be created with redirection of active input and output via pipes.  This amplifies and extends the utility of existing sensing and actuation approaches.  In addition to describing the design space of pipes, we also presented a design tool for the creation of tubes inside arbitrary 3D models and discuss many techniques that combine nicely wtih pipes.  We fabricated several example objects using this tool to explore new points in the design space.

We are heartened by our success: it opens new interactive device prototyping opportunities both for makers with consumer-grade 3D printers and research and industrial labs with higher-end machines.  We are excited to release our tool into the wild and see what people build.

### 2014-03-27 paper/designspace.tex
paper workin

Pipes can come in four types: open, return, semi-closed, and fully enclosed.  Each of these types offers distinct interaction capabilities (see Figure \ref{fig:pipespace}).  Since we are interested in interactive devices, we distinguish a ``user side" - part of the surface or interior of an object facing its user; and a ``system side" - part of the object (either interior or exterior) where other hardware such as electronics will be connected.  Definitionally, in the following paragraphs we refer to a pipe which is ``open on the user side'' to have its contents physically accessible to an interacting user at interaction time.  This includes, for example, pipes which function as fountains, but it does not include pipes whose ends have been capped to keep water inside (whether this capping was done at print time, or later as a manual modification).  Pipes whose contents are not physically accessible to the user are ``closed on the user side''.  A pipe which is ``open on the system side'' is physically open to the machine, and one which is ``closed on the system side'' has its contents physically cut off from the machine.  User and system manipulation of contents are possible with or without physical access, e.g., by resonating, pressurizing, or passing electricity through those contents. \bjoern{wordy paragraph}

@@

Open pipes originate from the system side and connect the user side, with both ends of the pipe open. This type of pipe may be used to create, for example, capacitive sensors: an open pipe filled with conductive paint can be connected to a sensing platform (e.g., Arduino) on the system side, while a user can touch the uncapped other end of the pipe.  Using Swept Frequency Capacitive Sensing \cite{Sato-touche} or other techniques, a user's touch of the open end can be sensed.  An open pipe can also be used for output, for example by creating in-air vortices as in \cite{Sodhi-aireal}.

A return pipe originates at the system and returns back to it, without offering physical access to the user.  By threading an electroluminescent (EL) wire through a clear return pipe, a maker can create a custom piece of neon art.  If a return pipe passes very close to the surface of a 3D printed model, warm or cold water passed through the pipe could be used for temperature-based haptic feedback.

Semi-closed pipes are open at the system end (for control of the enclosed medium) and closed at the user end.  The closed interface  may be fabricated from a malleable material: for example, a series of pipes terminated in thin rubber membranes on the user side can be actuated by an air pump to create haptic feedback, or a hinged cap can be blown open by air pressure.  Semi-closed pipes can also function as audio-generating resonance chambers.

A fully enclosed pipe has no openings on either the system side or the user side.  Fully enclosed pipes can be used as resonance chambers (e.g., for object identification), or as air bubbles (e.g., as used in \cite{Willis-printedoptics} for internal display).  It can also be used as a container for water or other particles that would otherwise fall out.

\bjoern{This section in general may need a small multipart figure that shows example photos or at least illustrations of these points in the design space. (Fig 2 is the abstract representation of the dimensions; here we have some concrete examples of points in the space).

Or maybe this comes later? it's a bit hard to follow these descriptions in the abstract.
}

``Gas'' comprises all compressible fluids.  Use of fluid pressure inside pipes can create haptic feedback at semi-closed interfaces, or the gases can be used as carriers for scents or fog.  As in \cite{Slyper-pressure}, structures can be engineered to change in air pressure when manipulated correctly (e.g., a spiral that changes pressure when twisted, but not pressed), and thus fluid pressure can also be used as an input.  \bjoern{Isn't this the same example as two paragraph above?}

Particulates, either printed in-place or inserted, can be of varying densities.  A single particle can be used for display.  Sparse particles in a stream of fluid can provide haptic feedback.  Dense particles in a semi-closed pipe allow for jamming-based interactions at any point on the surface of an object \cite{Follmer-jamming} \bjoern{may need more of an explanation.  you need jamming plus flexible surface material, right?}.

@@

\bjoern{Think of the ordering of dimensions is it a conscious choice to list this dimension last?}
Pipe designs may emphasize either their exterior features (connection points) or their interior paths.  These two features lead to different kinds of interfaces.  An example interface that focuses on the exterior connection points is the touch sensitive toy in Figure \ref{fig:toys}, where pipes must exit the toy at the frontal lobe, parietal lobes, etc.  An interface focused on the internal path of the pipe is the neon sign in Figure \ref{fig:neon}: output is based on the shape of the pipe.

>>>>

Pipes can come in four types: open, return, semi-closed, and fully enclosed.  Each of these types offers distinct interaction capabilities (see Figure \ref{fig:pipespace}).  Since we are interested in interactive devices, we distinguish a ``user side" - part of the surface or interior of an object facing its user; and a ``system side" - part of the object (either interior or exterior) where other hardware such as electronics will be connected.  Definitionally, in the following paragraphs we refer to a pipe which is ``open on the user side'' or ``open on the system side'' to have its contents \emph{physically accessible} to an interacting user or the system \emph{at interaction time}.  A pipe which is ``open on the user side'' or ``closed on the system side'' is physically cut off from the user or machine.  User and system manipulation of contents are possible with or without physical access, e.g., by resonating, pressurizing, or passing electricity through those contents.

@@

Open pipes (see Figure \ref{fig:pipespace}A) originate from the system side and connect the user side, with both ends of the pipe open. This type of pipe may be used to create, for example, capacitive sensors: an open pipe filled with conductive paint can be connected to a sensing platform (e.g., Arduino) on the system side, while a user can touch the uncapped other end of the pipe.  Using Swept Frequency Capacitive Sensing \cite{Sato-touche} or other techniques, a user's touch of the open end can be sensed.  An open pipe can also be used for output, for example by creating in-air vortices as in \cite{Sodhi-aireal}.

A return pipe (see Figure \ref{fig:pipespace}B) originates at the system and returns back to it, without offering physical access to the user.  By threading an electroluminescent (EL) wire through a clear return pipe, a maker can create a custom piece of neon art.  If a return pipe passes very close to the surface of a 3D printed model, warm or cold water passed through the pipe could be used for temperature-based haptic feedback.

Semi-closed pipes (see Figure \ref{fig:pipespace}C) are open at the system end (for control of the enclosed medium) and closed at the user end.  The closed interface  may be fabricated from a malleable material: for example, a series of pipes terminated in thin rubber membranes on the user side can be actuated by an air pump to create haptic feedback, or a hinged cap can be blown open by air pressure.  Semi-closed pipes can also function as audio-generating resonance chambers.

A fully enclosed pipe (see Figure \ref{fig:pipespace}D) is closed to both the system side and the user side.  Fully enclosed pipes can be used as resonance chambers (e.g., for object identification), or as air bubbles (e.g., as used in Printed Optics (\cite{Willis-printedoptics}) for internal display).  They can also be used as containers for water or other particles that would otherwise fall out.

Pipe designs may emphasize either their exterior features (connection points) or their interior paths.  These two features lead to different kinds of interfaces.  An example interface that focuses on the exterior connection points is the touch sensitive toy in Figure \ref{fig:toys}, where pipes must exit the toy at the frontal lobe, parietal lobes, etc.  An interface focused on the internal path of the pipe is the neon sign in Figure \ref{fig:neon}: output is based on the shape of the pipe.

@@

``Gas'' comprises all compressible fluids.  Fluid pressure inside semi-closed pipes can create haptic feedback, or gases can be used as carriers for scents or fog.  Slyper, et al.,  in \cite{Slyper-pressure} engineered structures to change in air pressure when manipulated correctly (e.g., a spiral that changes pressure when twisted, but not when pressed); fluid pressure can also be used as an input.

Particulates, either printed in-place or inserted, can be of varying densities.  A single particle can be used for display.  Sparse particles in a stream of fluid can provide haptic feedback.  Dense particles in a semi-closed pipe capped with flexible surface material allow for jamming-based interactions at any point on the surface of an object \cite{Follmer-jamming}; when fluid pressure in the pipe is reduced, the trapped particles jam to form a solid mass the same shape as their enclosing flexible membrane.

### 2014-03-27 paper/exampleobjects.tex
paper workin

To evaluate our tool's capabilities and to explore new points in the design space, we fabricated a set of five prototype objects designed using the tool.  All prototypes were fabricated in a single piece, unless described otherwise.

@@

We created a set of touch-sensitive toys and a companion app, reminiscent of the boat application in \cite{Harrison-acoustic}.  The brain, rabbit, and boat each have several distinct touch points on their surfaces.  These touch points are connected by an interior star topology of tubes filled with conductive copper paint, and touch sensing is performed via a single wire and SFCS (see Figure \ref{fig:toys}).  We built a smart base which can distinguish between the toys and also determine which toy is mounted: since each toy and each gesture has a distinct capacitive signature, we use a simple classifier trained to detect both toy and gesture based on profile.  All components were fabricated on a Makerbot, support-free.

@@

A custom radio built using tubes allows users to tune to different stations and listen in.  This device uses a network of disconnected open tubes filled with conductive paint to connect traditional electronic components (two potentiometers and a speaker) to a microcontroller (in our case, an Arduino).  We hid the microcontroller and a battery-driven power supply in the base of the radio to make it portable.  For the radio functionality, we attached an Si4703 to the Arduino.  We fabricated this device on our Makerbot, support-free but cut in two halves to allow the Arduino to be inserted.

@@

Our presence-aware pen holder can distinguish which tool or tools a user has picked up (see Figure \ref{fig:pens}).  Our pen holder uses a modification of the FlyEye technique described by Wimmer in \cite{Wimmer-flyeye} and contains open tubes filled with fiber optic cables, one per pen chamber.  We use a single cable per chamber; our 6mm diameter fiber optic cables, in comparison to the angel hair tubes used in the original work \bjoern{Mixed terminology between el wire and fiber optics? "angel hair" is a specific product name, not a general technical term (I think)}, can send and receive through the same cable.  At the base of each tube is a QRE1113 line sensor digital breakout board, which has an integrated IR emitter and receiver.   When a pen is in its appointed place, the emitted infrared light is reflected off its bottom and travels back to the receiver, where it registers as bright.  This prototype was built by our Objet and support was flushed post-print. \valkyrie{this whole paragraph is unclear}

@@

We created a maze game with a single particle trapped in a fully enclosed tube.  To fabricate the tube, we created it in two halves that were fastened together via glue, due to limitations in our ability to remove support material otherwise.  We created this prototype on our Makerbot in two halves to allow for insertion of the marble. \bjoern{You are giving two different reasons here for printing in pieces.}

@@

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/sign.png}
\caption{A two-state animated neon sign designed using our tool (a)(b) \valkyrie{Tovi: we could use your PDF hax to make this animate in place :)}.  (c) shows the SVG input files to create this sign, and (d) shows an isometric view of the tubes created by our tool. \valkyrie{d) is a lot less interesting if we are making planar tubes.  also, for c, should we show an overlay of the two states?  maybe in different colors?}}
\label{fig:neon}
\end{figure}

Neon art, perhaps best known for its association with Las Vegas, is traditionally made from hand-formed glass tubes containing neon gas.  The tubes light up when a current is passed through them.  For this type of art, the path of the tubes is of crucial importance, as it determines how the sign will look.  We designed a custom neon sign bearing the UIST logo and the waves of Hawaii.  The piece features open tubes due to support-flushing constraints, however the tubes could also be semi-closed (one end must remain open to connect the EL wire to its power source).  They have been threaded with EL wire which is lit in sequence to create an animation.  Our sign was fabricated on the Objet in two halves, cutting through the plane of the sign itself, with clear material for the sign around the input design and opaque everywhere else.  We manually flushed support from the non-planar connecting tubes.  The EL wire was threaded post-print, and is controlled by an EL wire sequencer.

>>>>

To evaluate and highlight our tool's capabilities, we fabricated a set of five prototype objects designed using the tool.  All prototypes were fabricated in a single piece, unless described otherwise.

\valkyrie{one way of combatting the enormousness of this paper is to put all these example objects figures into a single figure* instead}

@@

We created a set of touch-sensitive toys and a companion app, reminiscent of the boat application in Acoustic Barcodes \cite{Harrison-acoustic}.  The brain, rabbit, and boat each have several distinct touch points on their surfaces.  These touch points are connected by an interior star topology of tubes filled with conductive copper paint, and touch sensing is performed via a single wire and SFCS (see Figure \ref{fig:toys}).  We built a smart base which can distinguish between the toys and also determine which toy is mounted: since each toy and each gesture has a distinct capacitive signature, we use a simple classifier trained to detect both toy and gesture based on profile.

The models we used for the toys' exteriors were all downloaded from the free website Thingiverse.  All components were fabricated on a Makerbot, support-free.  After printing, we injected copper paint (CuPro-Cote) into the interior pipes using a craft syringe.  This paint requires approximately 2 days to fully dry in this configuration (3mm diameter tubes, maximum tube length 10cm).  Our smart base is powered by an Arduino Uno running code by Instructables users DZL and madshobye.

@@

A custom radio built using tubes allows users to tune to different stations and listen in.  This device uses a network of disconnected open tubes filled with conductive paint to connect traditional electronic components (two potentiometers and a speaker) to a microcontroller (in our case, an Arduino Pro Micro).  We hid the microcontroller and a battery-driven power supply in the base of the radio to make it portable.  For the radio functionality, we attached an Si4703 to the Arduino, which offers headphone-jack output.  We fabricated the radio itself on our Makerbot, support-free but cut in two pieces to allow the Arduino and battery pack to be inserted.

@@

Our presence-aware pen holder can distinguish which tool or tools a user has picked up (see Figure \ref{fig:pens}).  Our pen holder uses a modification of the FlyEye technique described by Wimmer in \cite{Wimmer-flyeye} and contains open tubes filled with fiber optic cables, one per pen chamber.  We use a single cable per chamber; our 6mm diameter fiber optic cables, in comparison to the fine tubes used in the original work , can send and receive through the same cable.  At the base of each tube is a QRE1113 line sensor digital breakout board, which has an integrated IR emitter and receiver.   When a pen is in its appointed place, the emitted infrared light is reflected off its bottom and travels back to the receiver, where it registers as bright.  This prototype was built by our Objet and support was flushed post-print. \valkyrie{this whole paragraph is unclear}

@@

We created a maze game with a single particle trapped in a fully enclosed tube.  This maze is based on an SVG we created by hand and processed with our internal path tube tool.  To fabricate the tube, we created it in two halves that were fastened together via glue, due to limitations in our ability to remove support material otherwise.  We created this prototype on our Makerbot in two halves, inserting a marble before closing.

@@

\begin{figure}[h!]
\centering
    \includegraphics[width=3.4in]{figures/sign.png}
\caption{A two-state animated neon sign designed using our tool (a)(b) \valkyrie{Tovi: we could use your PDF hax to make this animate in place :)}.  (c) shows the SVG input files to create this sign, and (d) shows an isometric view of the tubes created by our tool. \valkyrie{d) is a lot less interesting if we are making planar tubes.  also, for c, should we show an overlay of the two states?  maybe in different colors?}}
\label{fig:neon}
\end{figure}

Neon art, perhaps best known for its association with Las Vegas, is traditionally made from hand-formed glass tubes containing neon gas.  The tubes light up when a current is passed through them.  For this type of art, the path of the tubes is of crucial importance, as it determines how the sign will look.  We designed a custom neon robot head which can be animated to ``talk''.  The piece features semi-open pipes: one end must remain open to connect the EL wire to its power source.  The pipes have been threaded with EL wire which is lit in sequence to create an animation.  Our sign was fabricated on the Objet in two halves, cutting through the plane of the sign itself, with clear material for the sign around the input design and opaque everywhere else.  We manually flushed support from the non-planar connecting pipes.  The EL wire was threaded post-print, and is controlled by an EL wire sequencer.

### 2014-03-27 paper/limitations.tex
paper workin

\section{Limitations and Future Work}
Tubes, while a powerful tool for directing interaction to arbitrary locations on an object's surface, have their shortcomings.  Most obviously, their fundamental shortfall is that they cannot replace the active elements required for a functioning interactive device: they simply redirect sensing to different locations on an object's surface.  All tubes need to, at some point, connect to a sensor, microcontroller, or pump.

Some media are difficult to insert in tubes, or may not be compatible with all types of tubes.  Fluids in particular provide challenges.  For example, the conductive paint we used in our designs can take multiple days to dry in tubes of sufficient length.  Any liquid kept in tubes printed by the Objet can cause the exterior of the part to discolor due to seepage.  Hobbyist machines do not guarantee watertightness of their prints, so leaking of water or air can occur.

\valkyrie{Our design tool is not yet fully capable of realizing all possibilities in the design space of tubes.  It lacks the network manipulation to make mixing or splitting tubes, unless they are designed in an input vector graphics file.  Future work should also include assisting makers in the fabrication process itself by automatically cutting objects to reduce support material removal issues.  Additionally, as mentioned there are opportunities to improve the algorithm for neon sign routing, as ours does not guarantee a shortest-path routing. --------- this paragraph breaks the latex for some reason?}

>>>>

\section{Limitations and Future Work}
Pipes, while a powerful tool for directing interaction to arbitrary locations on an object's surface, have their shortcomings.  Most obviously, their fundamental shortfall is that they cannot replace the active elements required for a functioning interactive device: they simply redirect sensing to different locations on an object's surface.  All pipes need to, at some point, connect to a sensor, microcontroller, or pump.

Some media are difficult to insert in pipes, or may not be compatible with all types of pipes.  Fluids in particular provide challenges.  For example, the conductive paint we used in our designs can take multiple days to dry in tubes of sufficient length.  Any liquid kept in tubes printed by the Objet can cause the exterior of the part to discolor due to seepage.  Hobbyist machines do not guarantee watertightness of their prints, so leaking of water or air can occur.

Our design tool is not yet fully capable of realizing all possibilities in the design space of pipes.  It lacks the network manipulation to make mixing or splitting pipes, unless they are designed in an input vector graphics file.  Since pipes are committed to the mesh one at a time, our tool does not currently optimize multiple pipes at a time: we simply make greedy pairwise decisions.  This suggests the need for a better general 3D routing algorithm with multiple traces.  Future work should also include assisting makers in the fabrication process itself by automatically cutting objects to reduce support material removal issues.  Additionally, as mentioned there are opportunities to improve the algorithm for neon sign routing, as ours does not guarantee a shortest-path routing.
